---
title: "Data package for seafloor climage change (ArgentinaSSP245)"
author: "Chih-Lin Wei"
date: "`r Sys.Date()`"
output: 
  html_document: 
    smart: false
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  eval=FALSE)
```

# Installation

### To install the package, open R and type:

```{r}
install.packages("devtools")
```

### Then, you can install ArgentinaSSP245:

```{r}
devtools::install_github("chihlinwei/ArgentinaSSP245", dependencies = TRUE)
library(ArgentinaSSP245)
```

### License
GNU Affero General Public License v3.0

### Tutorials
* [Display seafloor climate change data](https://github.com/chihlinwei/ArgentinaSSP245/blob/main/tutorials/tute1.md)
* [Extract seafloor climate change data by polygon, polyline, or points](https://github.com/chihlinwei/ArgentinaSSP245/blob/main/tutorials/tute2.md)
* [Applying seafloor climate change data for habitat suitability modeling](https://github.com/chihlinwei/ArgentinaSSP245/blob/main/tutorials/tute3.md)
* [Model scenario comparisons](https://github.com/chihlinwei/ArgentinaSSP245/blob/main/tutorials/tute4.md)

### Citation

Chih-Lin Wei. (2024). chihlinwei/ArgentinaSSP245: Data package for seafloor climage change (ArgentinaSSP245) (v1.0.0). Zenodo. https://doi.org/10.5281/zenodo.13319179

